### 2020-04-30 Case Studies/03_spam_identification.rmd
spam identification case study
spam identification case study

library(codetools, quietly = TRUE, warn.conflicts = FALSE)
library(lattice, quietly = TRUE, warn.conflicts = FALSE)
library(fields, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)

@@

data.dir <- file.path(here::here(), "Case Studies", "datasets")

@@

>>>>

library(codetools, quietly = TRUE, warn.conflicts = FALSE)
library(lattice, quietly = TRUE, warn.conflicts = FALSE)
library(fields, quietly = TRUE, warn.conflicts = FALSE)
library(RColorBrewer, quietly = TRUE, warn.conflicts = FALSE)
library(tm, quietly = TRUE, warn.conflicts = FALSE)

@@

data.dir <- file.path(here::here(), "Case Studies", "datasets", "spam")

@@

# Using Statistics to Identify Spam

## Anatomy of an email Message

### Spam Data

```{r}
head(list.files(path = file.path(data.dir, "easy_ham")))
head(list.files(path = file.path(data.dir, "spam_2")))

directories <- paste(data.dir, list.files(data.dir), sep = .Platform$file.sep)

file_counts <- sapply(directories, function(dir) length(list.files(dir)))

total_files <- sum(file_counts)
total_files
```

```{r}
file_counts

idx <- c(1:5, 15, 27, 68, 69, 329, 404, 427, 516, 852, 971)

fn <- list.files(directories[1], full.names = T)[idx]

sampleEmail <- sapply(fn, readLines)
```

### Text Mining and Naive Bayes Classification

```{r}
msg <- sampleEmail[[1]]
which(msg == "")[1]

match("", msg)

splitPoint <- match("", msg)

msg[ (splitPoint - 2):(splitPoint + 6)]

header <- msg[1:(splitPoint - 1)]
body <- msg[ -(1:splitPoint) ]

```{r}

splitMessage <- function(msg) {
   splitPoint <- match("", msg)

   header <- msg[ 1:(splitPoint - 1)]
   body <- msg[ -(1:splitPoint)]

   return(list(header = header, body = body))
}

sampleSplit <- lapply(sampleEmail, splitMessage)

header <- sampleSplit[[1]]$header
grep("Content-Type", header)
grep("multi", tolower(header))

header[46]

headerList <- lapply(sampleSplit, function(msg) msg$header)

CTloc <- sapply(headerList, grep, pattern = "Content-Type")
CTloc

sapply(headerList, function(header) {
   CTloc <- grep("Content-Type", header)
   if( length(CTloc) == 0) return(NA)
   CTloc
})

hasAttach <- sapply(headerList, function(header) {
   CTloc <- grep("Content-Type", header)

   if(length(CTloc) == 0) return(F)
   grepl("multi", tolower(header[CTloc]))
})

hasAttach

header <- sampleSplit[[6]]$header
boundaryIdx <- grep("boundary=", header)
header[boundaryIdx]

sub(".*boundary=\"(.*)\";.*", "\\1", header[boundaryIdx])

header2 <- headerList[[9]]
boundaryIdx2 <- grep("boundary=", header2)
header2[boundaryIdx2]

sub('.*boundary="(.*)";.*', "\\1", header2[boundaryIdx2])

boundary2 <- gsub('"', "", header2[boundaryIdx2])

sub(".*boundary= *(.*);?.*", "\\1", boundary2)

boundary <- gsub('"', "", header[boundaryIdx])
sub(".*boundary= *(.*);?.*", "\\1", boundary)

```{r}
getBoundary <- function(header) {

   boundaryIdx <- grep("boundary=", header)
   boundary = gsub('"', "", header[boundaryIdx])
   gsub(".*boundary= *([^;]*);?.*", "\\1", boundary)
}

```{r}
sampleSplit[[6]]

boundary <- getBoundary(headerList[[15]])
body <- sampleSplit[[15]]$body

bString <- paste("--", boundary, sep = "")
bStringLocs <- which(bString == body)
bStringLocs

eString <- paste("--", boundary, "--", sep = "")
eStringLoc <- which(eString == body)
eStringLoc

msg <- body[ (bStringLocs[1] + 1) : (bStringLocs[2] - 1)]
tail(msg)

msg <- c(msg, body[ (eStringLoc + 1) : length(body) ])
tail(msg)
```

### Handle Attachments

```{r}

dropAttach <- function(email) {

}

### Extracting Words from the Message Body

```{r}
head(sampleSplit[[1]]$body)

msg <- sampleSplit[[3]]$body
head(msg)

### Stemming

```{r}
stopWords <- stopwords(kind = "en")
```

### Convert To Wordlist

```{r}
tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))

msg[ c(1, 3, 26, 27) ]

cleanMsg <- tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))
cleanMsg[ c(1, 3, 26, 27) ]

words <- unlist(strsplit(cleanMsg, "[[:blank:]]+"))

words <- words[ nchar(words) > 1 ]

words <- words[ ! (words %in% stopWords) ]

head(words)

```{r}

findMsgWords <- function(msg, stopwords = stopWords) {

   cleanMsg <- tolower(gsub("[[:punct:]0-9[:blank:]]+", " ", msg))

   words <- unlist(strsplit(cleanMsg, "[[:blank:]]+"))

   words <- words[ ! (words %in% stopwords) ]

   return(words)   
}

findMsgWords(msg)
```

### Prep Wrap-Up

```{r}
processAllWords <- function(dirName, stopWords) {
   # read all files in the directory
   fileNames <- list.files(dirName, full.names = T)

   # drop files that are not email, i.e., cmds
   notEmail <- grep("cmds$", fileNames)

   if( length(notEmail) > 0) fileNames <- fileNames[ -notEmail ]

   messages <- lapply(fileNames, readLines, encoding = "latin1")

   # split header and body
   emailSplit <- lapply(messages, splitMessage)

   # put body and header in own lists
   bodyList <- lapply(emailSplit, function(msg) msg$body)
   headerList <- lapply(emailSplit, function(msg) msg$header)
   rm(emailSplit)

   # determine which messages have attachments
   hasAttach <- sapply(headerList, function(header) {

      CTloc <- grep("Content-Type", header)

      if( length(CTloc) == 0) return(0)

      multi <- grep("multi", tolower(header[CTloc]))

      if( length(multi) == 0 ) return(0)

      multi
   })

   hasAttach <- which(hasAttach > 0)

   # drop attachments from message body
   bodyList[hasAttach] <- mapply(dropAttach, bodyList[hasAttach],
                                 boundaries, SIMPLIFY = F)

   # extract words from body
   msgWordsList <- lapply(bodyList, findMsgWords, stopWords)

   invisible(msgWordsList)
}
```

### Build Email Database

```{r}

msgWordList <- lapply(directories, processAllWords)
